---
title: "Stat 472 Calc Project"
author: "Emma, Jackson, Amanda, Megan"
date: "3/3/2020"
output: pdf_document
---

1. Data required for the Calculus retention study is in S1 File in the Supporting Information section of the paper. S2 File has information about the survey questions and response codings in the data set. Load the data and understand each column.


```{r echo=FALSE, message=FALSE}
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("knitr")
#install.packages("kableExtra")
install.packages("forcats")

library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(forcats)
```



```{r}
data <- read.csv("CSPCC_Switcher_Data.csv")
data$Q3FUS=paste(data$Q3FUS_No, data$Q3FUS_Yes)

```


# Switcher Coding

Switcher Groups

```{r}
s1 <- data %>% filter(Q26 == 1, Q3FUS_No == "No")
s2 <- data %>% filter(Q26 == 3, Q3FUS_No == "No")
s3 <- data %>% filter(is.na(Q26),Q5Post == 1, Q3FUS_No == "No")
s4 <- data %>% filter(is.na(Q26),Q5Post == 3, Q3FUS_No == "No")
s5 <- data %>% filter(Q26 == 1, Q3Post == 3, Q5Post == 1, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s6 <- data %>% filter(Q26 ==1, Q3Post == 2, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s7 <- data %>% filter(Q26 == 3, Q5Post == 1, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s8 <- data %>% filter(Q26 == 3, Q3Post == 2, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s9 <- data %>% filter(is.na(Q26), Q5Post == 1, Q3Post ==3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s10 <- data %>% filter(is.na(Q26), Q5Post ==1, Q3Post ==2, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
s11 <- data %>% filter(is.na(Q26), Q5Post == 3, Q3Post ==2, Q3FUS_No != "No", Q3FUS_Yes != "Yes")
```


Persister Groups

```{r}
p12 <- data %>% filter(Q26 == 1, Q3FUS_Yes == "Yes")

p13 <- data %>% filter(Q26 == 3, Q3FUS_Yes == "Yes")

p14 <- data %>% filter(is.na(Q26), Q5Post == 1, Q3FUS_Yes == "Yes")

p15 <- data %>% filter(is.na(Q26), Q5Post == 3, Q3FUS_Yes == "Yes")

p16 <- data %>% filter(Q26 == 1, Q3Post == 1, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p17 <- data %>% filter(Q26 == 1, Q5Post == 3, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p18 <- data %>% filter(Q26 == 1, Q5Post == 2, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes" )

p19 <- data %>% filter(Q26 == 1, is.na(Q5Post), Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p20 <- data %>% filter(Q26 == 3, Q3Post == 1, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p21 <- data %>% filter(Q26 == 3, Q5Post == 3, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")


p22 <- data %>% filter(Q26 == 3, Q5Post == 2, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p23 <- data %>% filter(Q26 == 3, is.na(Q5Post), Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes" ) 

p24 <- data %>% filter(is.na(Q26), Q5Post == 1, Q3Post == 1, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p25 <- data %>% filter(is.na(Q26), Q5Post == 3, Q3Post == 1, Q3FUS_No != "No", Q3FUS_Yes != "Yes")

p26 <- data %>% filter(is.na(Q26), Q5Post == 3, Q3Post == 3, Q3FUS_No != "No", Q3FUS_Yes != "Yes")


```

```{r}

switchdata <- rbind(s1,s2,s3,s4,s5,s6,s7,s8,s9,s10,
                    s11,p12,p13,p14,p15,p16,p17,p18,
                    p19,p20,p21,p22,p23,p24,p25,p26)

```

Table s1:

```{r}
x1=c("1", "2", "3", "4", "5", "6","7", "8", "9", "10", "11")
x2=c(count(s1), count(s2), count(s3), count(s4), count(s5), count(s6), count(s7), count(s8), count(s9), count(s10), count(s11))

x3 <- c("Y","M","NA","NA","Y","Y","M","M","NA","NA","NA")
x4 <- c("","","Y","M","Y","","Y","","Y","Y","M")
x5 <- c("","","","","M","N","M","N","M","N","N")
x6 <- c("N","N","N","N","NA","NA","NA","NA","NA","NA","NA")



tabS1 <- cbind(x1,x2,x3,x4,x5,x6)

colnames(tabS1) <- c("Switcher Group", "Number of Students","Beginning of term","End of term; reflect", "End of term", "Follow Up")
rownames(tabS1) <- NULL

kable(tabS1)


kable(tabS1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```
Table s1 (persister group)

```{r}
y1=c("12", "13", "14", "15", "16", "17", "18", 
     "19", "20", "21", "22", "23", "24", "25", "26")
y2=c(count(p12), count(p13), count(p14), count(p15), count(p16), 
     count(p17), count(p18), count(p19), count(p20), count(p21), 
     count(p22), count(p23), count(p24), count(p25), count(p26))
y3=c("Y", "M", "NA", "NA", "Y", "Y", "Y", "Y", "M", "M", "M", "M", "NA", "NA", "NA")
y4=c("", "", "Y", "M", "", "M", "N", "NA", "", "M", "N", "NA", "Y", "M", "M")
y5=c("", "", "", "", "Y", "M", "M", "M", "Y", "M", "M", "M", "Y", "Y", "M")
y6=c("Y", "Y", "Y", "Y", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA")

s1_p=cbind(y1, y2, y3, y4, y5, y6)

colnames(s1_p) <- c("Persister Group", "Number of Students","Beginning of term","End of term; reflect", "End of term", "Follow Up")
rownames(s1_p) <- NULL

kable(s1_p)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```



2. Prepare the data for analysis using S3 File. You should be left with 2,266 students (rows) after data preparation.

3. Reproduce the tables S1, S5, S6, S7, and S8.
Caution: There could be a typo in table S1.

```{r}
careerdata <- switchdata %>% 
  mutate(major = fct_collapse(factor(switchdata$Q60, levels=1:16), 
                              `1` = c("3","4","5","7","8","9"), 
                              `2` = "6", 
                              `3` = c("1","2"),
                              `4` = c("10","11","12","13","14","15"),
                              `5` = "16")) %>%
  filter(! is.na(major)) %>% ## filtering out NA's in major
  filter(! is.na(Q48)) %>% ## filtering out NA's in gender
  filter(! (is.na(Q3_SATMath) & is.na(Q7_ACTMath))) %>% ##std testing 
  filter(Q3_SATMath >= 200, Q3_SATMath <= 800, Q3_SATMath %% 10 ==0,  ) %>%
  filter(! (is.na(Q18Post_AskedQs) & is.na(Q18Post_Listened) & is.na(Q18Post_Applications) & is.na(Q18Post_Time) & is.na(Q18Post_ProblemSolver) & is.na(Q18Post_Explanations) & is.na(Q18Post_Appointments) & is.na(Q18Post_Discouraged))) %>% ## InstructorQuality 
filter(! (is.na(Q19Post_SpecificProblems) & is.na(Q19Post_WorkTogether) & is.na(Q19Post_Discussion) & is.na(Q19Post_Presentations) & is.na(Q19Post_Individually) & is.na(Q19Post_Lecture) & is.na(Q19Post_AskQuestions) & is.na(Q19Post_ExplainThinking))) %>%##StudentPractices
  filter(! is.na(Q18)) 
  

unique(careerdata$Q3_SATMath)

```


```{r}

IQ <- select(data, Q18Post_Applications, Q18Post_Appointments, Q18Post_AskedQs, Q18Post_Discouraged, Q18Post_Explanations, Q18Post_Listened, Q18Post_ProblemSolver, Q18Post_Time)

IQ$Q18Post_Discouraged <- 7-IQ$Q18Post_Discouraged

IQ <- na.omit(IQ)

pr_IQ <- princomp(IQ)

rescale_load_iq <- pr_IQ$loadings/sum(pr_IQ$loadings[,1])

round(rescale_load_iq[,1],3)

IQ <- IQ %>% mutate(Inst.Qual = (rescale_load_iq[,1]*(Q18Post_Applications+ Q18Post_Appointments+ Q18Post_AskedQs+ Q18Post_Discouraged+ Q18Post_Explanations+ Q18Post_Listened+ Q18Post_ProblemSolver+ Q18Post_Time))) 

```

```{r}

SC <- select(data, Q19Post_AskQuestions, Q19Post_Discussion, Q19Post_ExplainThinking, Q19Post_Individually, Q19Post_Lecture,Q19Post_Presentations,Q19Post_SpecificProblems, Q19Post_WorkTogether)

SC$Q19Post_Lecture <- 7 - SC$Q19Post_Lecture

SC <- na.omit(SC)

pr_SC <- princomp(SC)

round(pr_SC$loadings[,1],3)

rescale_load_sc <- pr_SC$loadings/sum(pr_SC$loadings[,1])

round(rescale_load_sc[,1],3)


SC <- SC %>% mutate(Student.Cent = (rescale_load_sc[,1]*(Q19Post_AskQuestions + Q19Post_Discussion + Q19Post_ExplainThinking + Q19Post_Individually + Q19Post_Lecture + Q19Post_Presentations + Q19Post_SpecificProblems + Q19Post_WorkTogether)))
```




